---
title: "Data tydying"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE}
library(dplyr)
library(tidyr)
library(readr)
```

# Useful shortcuts in R Markdown

ctrl alt i => short cut for R code chunck

show_col_types= FALSE (to omitt message description of data when using read_csv)

# Read in data

## Open data from url direclty

Data come from [Mike Byerly. Alaska commercial salmon catches by management region (1886- 1997). Gulf of Alaska Data Portal. df35b.304.2.](https://knb.ecoinformatics.org/view/df35b.304.2)

```{r}
catch_original<-read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")

head(catch_original)

```
# Clean and reshape data 

Drop notes and all column

```{r}
catch_data <- catch_original %>% 
  select(-All, -notesRegCode) 

```

Coerce Chinook column to be numeric instead of character

```{r}

catch_clean<-catch_data %>% 
  mutate(Chinook= if_else(Chinook=="I", "1", Chinook)) %>% 
  mutate(Chinook= as.numeric(Chinook))

head(catch_clean)

```

## Examine NA values

```{r}

i<-which(is.na(catch_clean$Chinook)) # pick up NA
i
catch_original[i,] #this line had a character (wrongly) instead of a number, so when we make is.numeric() above, it includes NA in this location.  

```


```{r}
catch_long<- catch_clean %>% 
  pivot_longer(cols = -c(Region, Year), names_to = "species", values_to = "catch")
```

```{r}
catch_long<-catch_long %>% 
  rename(catch_thousands = catch) %>% 
  mutate(catch = catch_thousands*1000) 
```

### Split-apply-combine

```{r}
mean_region<- catch_long %>% 
  group_by(Region) %>% 
  summarise(catch_mean = mean(catch), 
            n = n())#counts the number of rows within each group

mean_region
```

### Try out grouping and summarise

```{r}
catch_long

mean_year<-catch_long %>% 
  filter(Year>1881) %>% 
  group_by(Year, Region) %>% 
  summarise(mean_catch = mean(catch)) %>% 
  pivot_wider(id_cols = NULL, names_from = Region, values_from = mean_catch) %>% 
  arrange(Year) #ue a "-" before the -Year to order in descending order
#summarise(mean_catch = mean(catch), .groups = "drop") => to ungroup 

mean_year
```




```{r}


```

