---
title: "Exercise-tidying"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rvest) #read html, to get info from websites
library(readr)
library(dplyr)
library(janitor)
library(ggplot2)
```
```{r}

webpage <- read_html("https://www.pwrc.usgs.gov/bbl/manual/speclist.cfm")

tbls <- html_nodes(webpage, "table") %>% 
    html_table(fill = TRUE)

species <- tbls[[1]] %>% 
    clean_names() %>% #cleans up the spp names by replacing spaces by underscore, make everything lower case, etc 
    select(alpha_code, common_name) %>% 
    mutate(alpha_code = tolower(alpha_code))
head(species)
```


# 2.Read the following two files into your environment.


```{r}

pred <- read_csv("https://arcticdata.io/metacat/d1/mn/v2/object/urn%3Auuid%3A9ffec04c-7e2d-41dd-9e88-b6c2e8c4375e")
nest <- read_csv("https://arcticdata.io/metacat/d1/mn/v2/object/urn%3Auuid%3A982bd2fc-4edf-4da7-96ef-0d11b853102d")

head(pred)
head(nest)
```


# 3. Write a function that will translate species codes into common names.

```{r}

code_to_spp <- function(df1,species){
 merged <- left_join(df1, species, by=c("species" = "alpha_code")) 
   return(merged)
}

# make it general:
code_to_spp1 <- function(df1,df2,x,y){
 merged <- left_join(df1, df2, by=c(x = y)) 
   return(merged)
}


head(pred)

pred1 <- pred %>% 
      rename(alpha_code=species) %>% 
      left_join(., species, by=c("alpha_code"))

head(pred1)

# apply function:
pred_coded <- code_to_spp(pred, species)
head(pred_coded)


# tests
test <- code_to_spp(pred, species)

test1 <- code_to_spp1(pred, species, pred$species, species$alpha_code)

```


# 4. Calculate total number of predators by year and species, and plot the result.

```{r}
head(pred)

n.pred <- pred %>% 
  group_by(year, species) %>% 
  summarise(total = sum(count, na.rm=T))
  
names(n.pred)

ggplot(n.pred, mapping = aes(x = year, y = total, color = species)) +
  geom_line()

```

```


# 5. Calculate total number of eggs predated by year and species.

```{r}
head(nest)
head(pred)

n.eggs <- nest %>% 
  group_by(year, species) %>% 
  mutate(count = as.numeric(count))
  summarise(total = sum(count, na.rm=T))
  
names(n.pred)

ggplot(n.pred, mapping = aes(x = year, y = total, color = species)) +
  geom_line()

```






