---
title: "Untitled"
output: html_document
bibliography: references.bib
---
R code chunk that do not show any code (echo=FALSE) or show it (echo=TRUE)
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Second level header

### Third level header

Plain paragraph text

A new paragraph, this is not a new line.

#### Shopping list

**Fruits**

1.  apple
2.  *kiwi*
3.  banana

*latex equation* $\sum{x_n + 3}$

| Col1 | Col2 | Col3 |
|------|------|------|
|      |      |      |
|      |      |      |
|      |      |      |

: Caption table

*Insert citation* using the visual mode:

[@jakovac2021]

```{r}
#open an R code chunk

x<-2+2
x

```

cntrl + alt + i (pc): for including r chunk

cntrl + return (): for running

```{r}
x<-2*3
y<-x*3
y

```

```{r message=FALSE, warning=FALSE}
#Include all package dependencies needed for your code in the top of your document. This makes it easier to work collaboratively. 

library(readr)
library(dplyr)

```

# Introduction

-   read in data

```{r read-data, message=FALSE, warning=FALSE}
bg_chem<-read_csv("Data/BGchem2008data.csv", show_col_types = FALSE)
names(bg_chem)

```

# Analysis

-   calculate summary statistics
```{r}
nitrate<-mean(bg_chem$NO3)
nitrite<-mean(bg_chem$NO2)
amm<-mean(bg_chem$NH4)
phos<-mean(bg_chem$P)

```

-   calculate mean redfield ratio

```{r}
ratio<-(nitrate + nitrite+ amm)/phos
```

The calculated Redfield ratio is `r round(ratio,2)`


-   plot redfield ratio

```{r}
plot(bg_chem$P, bg_chem$NO2 + bg_chem$NO3 + bg_chem$NH4)
```

